\chapter{Building}
For basic information and some more see the file \file{Build.md} in the
\neotwo main directory.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Dependencies}
This section aims at listing the dependencies, and giving some
background on where there are used and how they are included in the
code.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{GSL/FGSL}
Then gnu scientific library is used via the fortran wrapper fgsl. The
fgsl in turn is wrapped by neo-2. Only those files which actually use
the fgsl code directly are compiled with the necessary include
information.
Currently besselfunction, numerical rootfinding, numerical integration
and bsplines are used by \neotwo.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{SuiteSparse}
Is used by sparse\_solver if \inputparameter{sparse\_solve\_method} is $1$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{SuperLU}
Is used by sparse\_solver if \inputparameter{sparse\_solve\_method} is
$2$ or $3$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{ATLAS}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{OpenBLAS}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{MPI/MyMPILib}
Used for parallelization. In some places also OpenMP is used.
Now included in libneo.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{HDF5/hdf5\_tools}
Used for input and output.
Now included in libneo.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{libneo}
Collection for common code of the work group.
Includes MyMPILib and hdf5\_tools.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{f90nml}
Python module, used to read and edit fortran namelist files.
If not present on your system you can install it with
\begin{verbatim}
pip3 install --local f90nml
\end{verbatim}
