\chapter{Programflow}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{PAR}

\paragraph{}
Set default namelist values.

\paragraph{}
Initialize hdf5.

\paragraph{}
Initialize MPI.

\paragraph{}
Read parameters from namelist files (\file{neo2.in}).

\paragraph{}
Check parameters.

\paragraph{}
If \codevariable{collop\_only\_precompute}, then construct/load the
collisionoperarot, otherwise continue.

\paragraph{}
Switch \codevariable{prop\_reconstruct}.

\paragraph{}
Deinitialize hdf5.

\paragraph{}
Deinitialize MPI.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{QL}

\paragraph{}
Set default namelist values.

\paragraph{}
Initialize hdf5.

\paragraph{}
Read parameters from namelist files (\file{neo2.in}).

\paragraph{}
Check parameters.

\paragraph{}
If the parameter say so, initialize a multispecies run and stop the
code.
Otherwise continue.

\paragraph{}
Initialize MPI.

\paragraph{}
Master writes version info.

\paragraph{}
If it is a reconstruction run call \codevariable{reconstruct\_prop\_dist}
and stop.

\paragraph{}
Call \codevariable{kin\_allocate}.

\paragraph{}
Call \codevariable{flint\_prepare} (setup magnetics).

Some things differ whether zylindrical or boozer coordinates are chosen.
The former has some additional lines for efit.
Differences in multispecies or not are due to different variables
(arrays vs. scalar).

This will read in the magnetic input file.
Efit files via compute\_aiota.
Boozer files via make\_mag\_device and stevvo\_x. Especially the boozer
case is hidden behind some interfaces.

Hierachy of the structures is as follows. At top is a \codevariable{device}.
From there it goes down to \codevariable{surface}, \codevariable{fieldline},
\codevariable{fieldperiod}, \codevariable{fieldpropagator} and
\codevariable{fieldripple}.

\paragraph{}
Compute/load collision operator.

\paragraph{}
Second preparation of flint.

\paragraph{}
Proptag related stuff?

\paragraph{}
Call \codevariable{flint}.

\paragraph{}
Unload collision operator.

\paragraph{}
Deinitialize MPI.

\paragraph{}
Deinitialize hdf5.
